---
title: "Gibbs"
author: "XC"
date: "20/07/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Appeals of Gibbs
- They gather most of their calibration from their target
- allow us to break complex problem (high-dim target such that rdwk HM almost impossible to build) into seriers of easier problems e.g. low-dim targets. 


## Two stage Gibbs sampler
It creates the joint distribution. 
If two rdv X, Y have joint distribution f(x, y), with conditional distribtion $f_{X|Y}, f_{Y|X}$, the the algo: 
- step 1: Set X0 = x0;
- For t = 1, 2, 3, ...
- step 2: $Yt \sim f_{Y|X}(.|x_{t-1})$
          $Xt \sim f_{X|Y}(.|y_t)$

Convergence of the MC is ensured unless the supports of the conditionals are _not connected_


### Example 7.2
A pair of distribution $X | \theta \sim Bin(n, \theta)$, $\theta \sim Beta(a, b)$,
leads to joint distribution $f(x, \theta) = f(x| \theta) \pi(\theta) = C_n^{x} \frac{\Gamma(a+b)}{\Gamma(a) \Gamma(b)} \theta^{x+a-} (1-\theta)^{n-x+b-1}$, so the posterior distribution of $\theta$, i.e. $f(\theta|x) \propto Beta (x+a, n-x+b)$

```{r}
Nsim <- 5000
n <- 15
a <- 3
b <- 7

X <- T <- array(0, dim = c(Nsim, 1))
T[1] <- rbeta(1, a, b)
X[1] <- rbinom(1, n, T[1])

for (i in 2:Nsim) {
  X[i] <- rnorm(1, n, T[i-1])
  T[i] <- rbeta(1, X[i] + a, n - X[i] + b)
}
```

```{r}
par(mfrow = c(1, 2))
hist(X,  xlab = "X", freq = F, ylab = "marginal density")
lines(density(X), col = "gold", lwd = 2)

hist(T, freq = F, xlab = expression(theta), ylab = "marginal density")
lines(density(T), col = "gold", lwd = 2)
```



Remark: 
  - this is a special case as the marginal or posterior of theta is available in _closed_ form. 
  
  

### Example 7.3

```{r}
x <- c(91,504,557,609,693,727,764,803,857,929,970,1043,1089,1195,1384,1713)  # data

a <- 3; b <- 3; theta_0 <- 5; tau2 <- 10   # specified values for hyper parameters

xbar <- mean(x)
n <-length(x)
Nsim <- 10000

sigma2 <- theta <- rep(0, Nsim)                     # inti arrays

sigma2[1] <- 1 / rgamma(1, shape = a, rate = b)     # initializae sigma2
B <- sigma2[1] / (sigma2[1] + n * tau2) 
theta[1] <- rnorm(1, mean = B * theta_0 + (1 - B) * xbar, sd = sqrt(B * tau2)) # init theta

for (i in 2:Nsim) {
  B <- sigma2[i - 1] / (sigma2[i - 1] + n * tau2)
  theta[i] <- rnorm(1, mean = B * theta_0 + (1 - B) * xbar, sd = sqrt(B * tau2))
  Ra <- sum((x - theta[i])^2)
  sigma2[i] <- 1 / rgamma(1, shape = n / 2 + a, rate = 0.5 * Ra + b)
}

```

```{r}
log(xbar)   # 6.769068

sd(x)       # 378

log(sd(x))

```
Now check the posterior mean of theta and sigma2
```{r}
mean(theta)
mean(sqrt(sigma2))

mean(log(sqrt(sigma2)))   # 6.707253


par(mfrow = c(1, 2))

hist(log(theta), breaks = 150)
hist(log(sqrt(sigma2)), breaks = 150)
#hist(log(theta), ylab = "posterior density of log(theta)")
#hist(log(sqrt(sigma2)), ylab = "posteior density of log(sigma)")
```




