---
title: "Metropolis with Gibbs and hybrid"
author: "XC"
date: "23/07/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hybrid

When the full conditional distribution of the gibbs is not able to be simulated from the rd number generator, e.g, within a set of full conditional densities f1, ..., fp, fi is not conventional, then _Metropolis within gibbs_ can be adopted. 


In stead of simulating $X_i \sim f_i(x_i | x_1^{(t+1)}, x_2^{(t+1)}, ..., x_{i-1}^{(t+1)}, x_{i+1}^{(t)}, x_{i+2}^{(t)})$, we can simulate one single step of _any_ MCMC scheme associated with the stationary distribution $f_i(x_i | x_1^{(t+1)}, x_2^{(t+1)}, ..., x_{i-1}^{(t+1)}, x_{i+1}^{(t)}, x_{i+2}^{(t)})$

  - a simple resolution is use a rd walk metropolis algo centered at $x_i^{(t)}$
  - the validility of the resulting algo is the joint distribution f remains the stationary distribution of the correspoding Markov chain. 
  - not using pure metropolis-hasting is because designing a MH on a large-dimension of target is challenging or even impossible. 
  - The fundamental gain in using Gibbs-like structure is it breaks down a complex model into a large number of smaller/simpler targets, where local MH algo can be desinate
  


## Example 7.18  
```{r}
Nsim <- 500
X <- rep(rexp(1), Nsim)
Y <- rep(rexp(1), Nsim)

for (i in 2:Nsim) {
  X[i] <- rexp(1, rate = Y[i-1])
  Y[i] <- rexp(1, rate = X[i])
}

par(mfrow = c(1, 2))
hist(X)
plot(cumsum(X) / (1:Nsim))
  
```
  
Remark
  - Convergence issue
  - due to these conditional distr do not correspond to any joint distribution
  - the only possible joint dist is $f(x, y) = exp(-xy)$, but has no finite integral
  

So the only way to make sure the Gibbs sampler you're using is valid is to check the joint distribution has a fintie integral. 

## Example 7.19

If improper prior are used in Gibbs, then the posterior must always be checked for proprioety. 
And it's always the case that improper prior on the variance is more problematic than on the mean. 




















